Constant-round Blind Classical Verification of Quantum Sampling

### 2020-04-12 hamiltonian.tex
Updates from Overleaf

To deal with $\SampBQP$, we will still employ $\Hin$ and $\Hprop$ to certify the circuit's input and gate-by-gate evolution. However, in this case, we care about the entire final state of the circuit, rather than the energy on the output qubit. The history state is still in the ground space of $\Hin + \Hprop$.  
One approach to certify the entire final state encoded inside the history state is to make sure that the history state is the unique ground state of $\Hin + \Hprop$ and all other orthogonal states will have much higher energies. 
Namely, we need to construct some $\Hin'+ \Hprop'$ with the history state as the unique ground state and with a large \emph{spectral} gap between the ground energy and excited energies. 
This guarantees that any state with close-to-ground energy must also be close to the history state. 
We remark that this is a much stronger and different requirement from most local Hamiltonian constructions where the focus is on the ground energy. 
However, we show it can indeed be achieved by using the \emph{perturbation} technique 
developed in the local Hamiltonian literature (e.g.~\cite{kempe_kitaev_regev_2006}) for a different purpose.

@@

We begin by recalling the form that local Hamiltonian certificates:

\begin{dfn}
	\label{dfn:groundstate}	
	Let $C=U_T\ldots U_1$ be a quantum circuit consisting of $T$ gates, and let $x\in\set{0,1}^n$. Then we denote
	$$\ket{\psi_{C(x)}}=\sum_{t=0}^TU_t\ldots U_1\ket{x}\otimes\ket{\hat{t}}$$
	We call the first register of $\ket{\psi_{C(x)}}$ the data register, and the second register of $\ket{\psi_{C(x)}}$ the time register.
\end{dfn}

\begin{rmk}
	\label{idpadding}
	One can first pad $C$ with identity gates at the end, doubling its number of gates and obtaining another circuit $C'$.
	Then by measuring the time register of $\ket{\psi_{C'(x)}}$ one obtains with probability $\frac{1}{2}$ the state $C(x)$.
\end{rmk}

@@

@@

The second one is a way to verify ground states of local Hamiltonians, again with some extra properties:
\begin{thm}
	\label{ThmXZCheck}
	Given circuit $C$ and an input string $x\in\set{0,1}^*$, there exists a $\BQP$ algorithm that accepts $\ket{\phi}$ with probability $\frac{1}{2}-\Omega(\frac{1}{\poly(T)})\braket{\phi|H_{C(x)}|\phi}$. (In particular, it accepts the $\ket{\psi_{C(x)}}$ with probability $\fot$.) Furthermore, this algorithm doesn't apply any quantum gates, using only $X$ and $Z$ measurements.
\end{thm}

\subsection{Circuits of Toffoli and Hadamard gates}

Old notation: Let $\Lambda_c(U)$ denote the gate $U$ controlled on qubit $c$, and $\Lambda^2_{f, s}(U)$ denote the gate $U$ controlled by both $f$ and $s$. I.e. $\Lambda^2_{1, 2}(X_3)$ would be a Toffoli gate.
\Ethan{TODO: replace with something that looks better and doesn't conflict with the use of $\Lambda$ as measurement.}

Now we present the proof that $\set{H, \Lambda^2(X)}$ is indeed a universal gate set from \cite{quant-ph/0301040}: \Ethan{This definitely belongs in the appendix instead...}
\begin{thm}
	Let $C$ be a circuit that:
	\begin{itemize}
		\item consists of $T$ gates, each either $H$ and $\Lambda(P(i))$
		\item uses $n$ qubits
	\end{itemize}
	Then a classical machine given $C$ can compute a circuit $C'$ that:
	\begin{itemize}
		\item consists of at most $4T$ gates, each either $H$ or $\Lambda^2(X)$ \hannote{don't like this notation}.
		\item uses $n+1$ qubits
		\item Let $x\in\set{0,1}^n$. Let $x|| 0\in\set{0,1}^{n+1}$ be $x$ concatenated with $0$.
		The measurement result of a tensor product of $X$ and $Z$ operators on the first $n$ qubits of $C'(x||0)$ has the same distribution as that on $C(x)$.
	\end{itemize}
\end{thm}
\begin{prf}
	We know that Hadamard gate and controlled phase gate is universal from \cite{kitaev_1997}
	$$H=\frac{1}{\sqrt{2}}\begin{pmatrix}1&1\\1&-1\end{pmatrix}$$
		$$\Lambda(P(i))=\begin{pmatrix}1&0&0&0\\0&1&0&0\\0&0&1&0\\0&0&0&i\end{pmatrix}$$

	Now consider the transform on quantum states
	$$\mathcal{F}(\ket{\phi})=(\Re\ket{\phi})\otimes\ket{0}+(\Im\ket{\phi})\otimes\ket{1}$$
	where $\Re$ and $\Im$ denote real and imaginary parts respectively.

	This transform commutes with Hadamard gates on the respective qubit. On the other hand, exchanging $\mathcal{F}$ with a controlled phase gate turn it into a combination of Hadamard and Toffoli gates. Mathematically,
	$$\mathcal{F}\circ H_s=H_s\circ\mathcal{F}$$
	$$\mathcal{F}\circ\Lambda_f(P(i)_s)=\Lambda^2_{f,s}(X_0Z_0)\circ\mathcal{F}=\Lambda^2_{f,s}(X_0)H_0\Lambda^2_{f,s}(X_0)H_0\circ\mathcal{F}$$

	We construct $C'$ so that $\mathcal{F}\circ C=C'\circ\mathcal{F}$ by following the computation above. It satisfies the required properties by construction.
	\begin{itemize}
		\item $C'$ uses only Hadamard and Toffoli gates.
		\item Exchanging $\mathcal{F}$ with $H$ doesn't change the circuit size. Exchanging $\mathcal{F}$ with controlled phase gate turns it into $4$ gates. So the final result is at most $4T$ gates.
		\item When $x$ is classical, $\mathcal{F}(x)=x||0$. So $F\circ C(x)=C'(x||0)$. It is simple \hannote{need a little more work?}to verify that $\mathcal{F}$ preserves $X$ and $Z$ measurement results on the first $n$ qubits.
	\end{itemize}
\end{prf}

\subsection{Reducing quantum circuit to local Hamiltonian}

@@

\subsection{Checking the ground state of the local Hamiltonian}

>>>>

To deal with $\SampBQP$, we will still employ $\Hin$ and $\Hprop$ to certify the circuit's input and gate-by-gate evolution. However, in this case, we care about the entire final state of the circuit, rather than the energy on the output qubit. The history state remains in the ground space of $\Hin + \Hprop$.  
Our approach to certify the entire final state (which is encoded inside the history state) is to make sure that the history state is the unique ground state of $\Hin + \Hprop$ and all other orthogonal states will have much higher energies. 
Namely, we need to construct some $\Hin'+ \Hprop'$ with the history state as the unique ground state and with a large \emph{spectral} gap between the ground energy and excited energies. 
It is guaranteed that any state with close-to-ground energy must also be close to the history state. 
We remark that this is an additional strong requirement, whereas most local Hamiltonian constructions focus on the ground energy. 
We show it can indeed be achieved by using the \emph{perturbation} technique 
developed in the local Hamiltonian literature (e.g.~\cite{kempe_kitaev_regev_2006}) for a different purpose. 
Another example of local Hamiltonian construction focusing on the spectral gap can be found in~\cite{adiabatic} where the purpose is to simulate quantum circuits by adiabatic quantum computation.

We need, however, two more further twists to serve our purpose. 
First, as we will eventually measure the final state in order to obtain classical samples, we need that the final state occupies a large fraction of the history state. We can simply add dummy identity gates for that purpose. 
Second, as we are only able to perform $X$ or $Z$ measurement by techniques from~\cite{FOCS:Mahadev18a},
we need to construct X-Z only local Hamiltonians. 
\XW{TOADD: some old references}
We achieve this by choosing one particular universal gate set and will include our construction for the sake of completeness.

\XW{TOADD: some orgainization}

@@

\subsection{Specific Construction}

Recall the definition of the history state which serves as a transcript of the circuit evolution~\cite{kitaev2002classical}:

\begin{dfn}[History-state]
	\label{dfn:groundstate}	
    Given any quantum circuit $C=U_T\ldots U_1$ of $T$ elementary gates and input $x\in\{0,1\}^n$, the \emph{history}-state $\histpsi{C(x)}$ is defined by
    \begin{equation}
        \histpsi{C(x)} \equiv \frac{1}{\sqrt{T}}\sum_{t=0}^TU_t\ldots U_1\ket{x,0}\otimes\ket{\hat{t}},
    \end{equation}
    where the first register refers to the input, the second refers to the work space which is initialized to $\ket{0}$, and the last refers to the clock space which encodes the time information. Note that $\hat{t}$ could be some representation of $t$=0,..., $T$ up to design choice. 
\end{dfn}

\XW{define $k$-local XZ Hamiltonian}

@@

\XW{cite the results that Toffoli and Hadamard gates
and add the fact}

\XW{main construction and proofs}

@@

%\subsection{Circuits of Toffoli and Hadamard gates}

% Old notation: Let $\Lambda_c(U)$ denote the gate $U$ controlled on qubit $c$, and $\Lambda^2_{f, s}(U)$ denote the gate $U$ controlled by both $f$ and $s$. I.e. $\Lambda^2_{1, 2}(X_3)$ would be a Toffoli gate.
% \Ethan{TODO: replace with something that looks better and doesn't conflict with the use of $\Lambda$ as measurement.}

@@

\subsection{TBA}

### 2020-04-12 qpip0.tex
Updates from Overleaf

In her groundbreaking work, Mahadev~\cite{FOCS:Mahadev18a} gave the following 4-round interactive protocol to let a $\BQP$ machine "commit a $XZ$ measurement" to a classical machine.

@@

\begin{lem}~\label{lem:trivial-4-round-strategy}
    For any input, there exist a prover strategy for Protocol~\ref{proto:urmila4} that is accepted with probability $1-\negl{n}$ in the testing round and \hannote{binding?}.
\end{lem}

@@

\begin{lem}[another triangle inequality for computational distance]\label{lem:computational-triangle}
Let $D_1$, $D_2$ be two distributions such that $D_1=D_3+D_4$ and $D_2=D_5+D_6$, where $D_3$, $D_4$, $D_5$, and $D_6$ are subnormalized distributions.

@@

	\item The verifier and prover run Protocol~\ref{proto:urmila4}, with the following parameters: $h$ is chosen from previous step, $n_{commit}=M \times$ (number of qubits in $\ket{\psi_{C'(x)}}$), where $n_{commit}$ is the $n$ for Protocol~\ref{proto:urmila4}, and the honest prover choose $M$ copies of $\ket{\psi_{C'(x)}}$ as $\rho$.

@@

$$z_i\defeq \sum_z \proj{z} \cdot \vev{e_i,\psi|U^\dag P_{acc,i,z} U|e_i,\psi},$$
where $\ket{e_i,\psi}=\ket{e_i}_\regC\ket{\psi}_{\regX,\regZ}$ and $U$ is the unitary the prover applies on the last round.

@@

For every $i\in [m]$ and every prover strategy $(U_0,U)$ for Protocol~\ref{proto:QPIP0samp}, consider the following composite strategy of the prover for the naive $\QPIP_0$ Protocol, Protocol~\ref{proto:qpip0_naive}: The prover tries to run $U_0$ by taking the verifier's input as $i$-th copy of protocol~\ref{proto:QPIP0samp} and simulating other $m-1$ copies by himself. The prover then \hannote{picks a uniformly random $\gamma$ ?}  tries to generate $\ket{\psi_{1^{i-1}1,\gamma}}$ by applying $G_{i,1,\gamma}G_{i-1,1,\gamma} \cdots G_{2,1,\gamma}G_{1,1,\gamma}$. If the prover fails to generate $\ket{\psi_{1^{i-1}1,\gamma}}$, he throws out everything and run $U_{trivial,0}$ described  in  Lemma~\ref{lem:trivial-4-round-strategy} instead. This is possible because all the inputs are classical. On the fourth round, if the prover had generated $\ket{\psi_{1^{i-1}1,\gamma}}$, he reply with the $i$-th register of $\ext_i\left(\frac{\ket{e_j}_{\regC}\ket{\psi_{1,i,\gamma}}_{\regX,\regZ}}{\|\ket{\psi_1}_{\regX,\regZ}\|}\right)$, where $\ext_i$ is specified in property~\ref{property:partition-binding} of Lemma~\ref{lem:partition2}. otherwise he continues to run the strategy of Lemma~\ref{lem:trivial-4-round-strategy}. We denote the random variable the verifier would get on the testing round as $z_{construct}$.

>>>>

In her groundbreaking work, Mahadev~\cite{FOCS:Mahadev18a} gave a 4-round interactive protocol to let a $\BQP$ machine "commit a $XZ$ measurement" to a classical machine, which we describe below while omitting the implementation details.

@@

\begin{lem}[\hannote{ref}\hannote{unused}]~\label{lem:trivial-4-round-strategy}
    For any input, there exist a prover strategy for Protocol~\ref{proto:urmila4} that is accepted with probability $1-\negl{n}$ in the testing round and \hannote{binding?}.
\end{lem}

@@

\begin{lem}[another triangle inequality for computational distance \hannote{unused}]\label{lem:computational-triangle}
Let $D_1$, $D_2$ be two distributions such that $D_1=D_3+D_4$ and $D_2=D_5+D_6$, where $D_3$, $D_4$, $D_5$, and $D_6$ are subnormalized distributions.

@@

	\item The verifier and prover run Protocol~\ref{proto:urmila4}, with the following parameters: $h$ is chosen from previous step, $n_{commit}=M \times$ (number of qubits in $\ket{\psi_{C'(x)}}$), where $n_{commit}$ is the $n$ for Protocol~\ref{proto:urmila4}, and the honest prover choose $M$ copies of $\ket{\psi_{C'(x)}}$ as $\rho$. We also allow a cheating prover to send $\bot$ in the second round of Protocol~\ref{proto:urmila4} to abort the process.\footnote{This is useless for the cheating prover but useful for proving our theorem.} If the verifier receives $\bot$ in the second round, he outputs $(\Rej,\bot)$ and terminates the protocol.

@@

$$z_i\defeq \sum_z \proj{z} \cdot \vev{e_i,\psi|U^\dag P_{acc,i,z} U|e_i,\psi},$$
where $\ket{e_i,\psi}=\ket{e_i}_\regC\ket{\psi}_{\regX,\regZ}$ and $U$ is the unitary the prover applies on the last round. \hannote{need to average $pk$ here.}

@@

For every $i\in [m]$ and every prover strategy $(U_0,U)$ for Protocol~\ref{proto:QPIP0samp}, consider the following composite strategy of the prover for the naive $\QPIP_0$ Protocol, Protocol~\ref{proto:qpip0_naive}: The prover tries to run $U_0$ by taking the verifier's input as $i$-th copy of Protocol~\ref{proto:urmila4} in Protocol~\ref{proto:QPIP0samp} and simulating other $m-1$ copies by himself. The prover then \hannote{picks a uniformly random $\gamma$ ?}  tries to generate $\ket{\psi_{1^{i-1}1,\gamma}}$ by applying $G_{i,1,\gamma}G_{i-1,1,\gamma} \cdots G_{2,1,\gamma}G_{1,1,\gamma}$. If the prover fails to generate $\ket{\psi_{1^{i-1}1,\gamma}}$, he throws out everything and run $U_{trivial,0}$ described  in  Lemma~\ref{lem:trivial-4-round-strategy} instead. This is possible because all the inputs are classical. On the fourth round, if the prover had generated $\ket{\psi_{1^{i-1}1,\gamma}}$, he reply with the $i$-th register of $\ext_i\left(\frac{\ket{e_j}_{\regC}\ket{\psi_{1,i,\gamma}}_{\regX,\regZ}}{\|\ket{\psi_1}_{\regX,\regZ}\|}\right)$, where $\ext_i$ is specified in property~\ref{property:partition-binding} of Lemma~\ref{lem:partition2}. otherwise he continues to run the strategy of Lemma~\ref{lem:trivial-4-round-strategy}. We denote the random variable the verifier would get on the testing round as $z_{construct}$.
